---
title: "ODEsensitivity"
author: "Dirk Surmann"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ODEsensitivity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  , comment = "#>"
  , fig.width = 7
)
library(ODEsensitivity)
```

## Introduction
The goal of sensitivity analysis is to examine how sensitive a mathematical model responds to variations in its input variables.
Here we focus on the sensitivity analysis of ordinary differential equation (ODE) models via Morris screening.

If the assumption of a uniform distribution on the domain intervals doesn't hold, the Morris screening method cannot be used and the variance-based Sobol' method should be considered instead.
In this case, simply switch from using the function `ODEmorris` to the function `ODEsobol`.

## Analyse the Lotka-Volterra Equations
The Lotka-Volterra equations describe a predator and its prey's population development and go back to Lotka (1925) and Volterra (1926).
The prey's population at time $t$ (in days) will be denoted with $P(t)$ and the predator's (or rather consumer's) population with $C(t)$.
$P(t)$ and $C(t)$ are called state variables.
This ODE model is two-dimensional, but it should be noted that ODE models of arbitrary dimensions (including one-dimensional ODE models) can be analyzed with ODEsensitivity.

### Model Definition
Now we define the model according to the definition in `deSolve::ode()`.
```{r}
LVmod = function(Time, State, Pars) {
  with(as.list(c(State, Pars)), {
    Ingestion <- rIng * Prey * Predator
    GrowthPrey <- rGrow * Prey * (1 - Prey/K)
    MortPredator <- rMort * Predator
    dPrey <- GrowthPrey - Ingestion
    dPredator <- Ingestion * assEff - MortPredator
    return(list(c(dPrey, dPredator)))
  })
}
```
Each of the five parameter names, their lower and upper boundaries, the initial values for the state variables
and the timepoints of interest are saved in separate vectors:
```{r}
LVpars = c("rIng", "rGrow", "rMort", "assEff", "K")
LVbinf = c(0.05, 0.05, 0.05, 0.05, 1)
LVbsup = c(1.00, 3.00, 0.95, 0.95, 20)
LVinit = c(Prey = 1, Predator = 2)
LVtimes = c(0.01, seq(1, 50, by = 1))
```

### Sensitivity Analysis
The sensitivity analysis of a general ODE model can be performed by using the generic function `ODEsensitivity::ODEmorris()`.
```{r}
set.seed(1618)
LVres_morris = ODEmorris(mod = LVmod, pars = LVpars, state_init = LVinit
                         , times = LVtimes, binf = LVbinf, bsup = LVbsup
                         )
```
Let's take a look at the output `LVres_morris`.
```{r}
str(LVres_morris, give.attr = FALSE)
```
The first row of each state variable contains a copy of all timepoints.
The other rows contain the Morris sensitivity indices $\mu$, $\mu^\star$, and $\sigma$ for all 5 parameters and all 51 timepoints.

### Plotting
ODEsensitivity provides a `plot()` method for objects of class `ODEmorris`:
```{r}
plot(LVres_morris)
```
`plot.ODEmorris()` has two important arguments: `pars_plot` and `state_plot`.
Using `pars_plot`, a subset of the parameters included in the sensitivity analysis can be selected for plotting (the default is to use all parameters).
`state_plot` gives the name of the state variable for which the sensitivity indices shall be plotted (the default being the first state variable):
```{r}
plot(LVres_morris, pars_plot = c("rIng", "rMort", "assEff"), state_plot = "Predator")
```

